\chapter{Decisions}


\section{Why behaviour-driven-development BDD?}

\begin{itemize}
\item Own experience from working with BDD out in the field.

\item Dan North, Introduction to BDD, https://dannorth.net/introducing-bdd/

\paragraph{}
    >If we could develop a consistent vocabulary for analysts, testers, developers, and the business, then we would be well on the way to eliminating some of the ambiguity and miscommunication that occur when technical people talk to business people.
    
\paragraph{}
    >A story’s behaviour is simply its acceptance criteria – if the system fulfills all the acceptance criteria, it’s behaving correctly; if it doesn’t, it isn’t. So we created a template to capture a story’s acceptance criteria.
\end{itemize}

\begin{lstlisting}[language=text,caption={As a, I want, so what}]
    As a [X]
    I want [Y]
    so that [Z]
\end{lstlisting}

\begin{lstlisting}[language=text,caption={given, when, then}]
    Given some initial context (the givens),
    When an event occurs,
    then ensure some outcomes.
\end{lstlisting}



\section{Why Rust?}
\begin{itemize}
\item Multiple libraries in the SSI community written in Rust already
\item The only official implementation of DIDComm v2 is in Rust
\item Strong typing
\item Safety/security features eliminates whole classes of bugs:
    \begin{itemize}
    \item Memory safety - allocating/freeing
    \item Boundary safety
    \item Null pointer safety
    \end{itemize}
\item Compiler catches many bugs before run-time
\item Modern language from this century
\item Performance like C
\item Backed by industry champs like Mozilla, Microsoft, Linux Foundation
\end{itemize}



\section{Why a command line interface - no GUI?}

\begin{itemize}
\item GUI does not solve our problem statement any better.
\item A degree in Programming, does not get extra points for design.
\item Time considered better spent on lower-level stuff.
\end{itemize}



\section{Why only support a single cryptographic toolkit?}
\begin{itemize}

    \item Why only support ed25519/x25519 for signing and encryption?
    \item To limit scope.
    \item Supporting multiple toolkits, does not solve our problem-statement any better.
    \item One is enough to prove the point.
    \item One could rewrite the software to support multiple cryptographic toolkits later, using a plugin-based architecture.
\end{itemize}



\section{Why only support a single DIDComm transport?}
\begin{itemize}
    \item To limit scope.
\end{itemize}
