\chapter{Reflection}

\section{Joining a large existing open-source community as a complete noob}

I spent some time in the beginning to get in touch with community experts. This resulted in me joining the DIF foundation, to get access to it's Slack-channel, and ask direct questions to the people that know the most about SSI, and it's development, in the world.

\section{Learning a new low-level language - Rust}

I had a lot of experience from programming in low-level languages and was surprised on how different it felt to program in Rust compared to C, C++, which are very similar to each other. Rust shows that inovation is still possible in this area, after many years with most innovation happening in higher-level languages.

\section{Techniques to overcome writers block}
\begin{itemize}
    \item Write something every day, no exceptions. Even if it just a single word. Do not stop ever.
    \item Allow yourself to write stupid things, and then rewrite. Rewriting is easier than writing.
    \item Write bullet points.
    \item Embrace chaos.
    \item Don't be afraid to delete and move things around.
\end{itemize}



\newpage

\section{How to talk to regular people about Self-Sovereign Identity?}

A list of concepts regular-people understand:
\begin{itemize}
    \item Attester
    \item Vitnemål
    \item Kursbevis
    \item Kontrakter
    \item ID
    \item Skjøter
    \item Titler
    \item Fødselsattest
    \item Pass
    \item BankID
    \item Førerkort
    \item Bankkontoer
    \item Lånebevis
    \item Penger (cryptocurrency)
    \item Helsejournal - 200 different countries, 200 different helth care apps?
    \item Resepter
    \item ...and more
\end{itemize}
